Missing data: Current practice in football research and recommendations for reporting missingness

### 2020-09-02 paper/paper.Rmd
some more expansion on MCAR and friends

For a value to be missing, it should have been observed in the data, but was not. For example, a goal scored in football takes place, but the scorekeeper wasn't paying attention and did not record an event. This "goal" is now a missing value. Compare this to a game being rained out and no game taking place. There is no data to be observed, so there is no missing data.

Data does not go missing the same way every time. There are three broad
categorisations describing why data can be missing: Missing Completely
at Random (MCAR), Missing at Random (MAR), and Missing Not at Random
(MNAR). These categorisations help describe whether the missingness
occurs by chance (MCAR), due to some observed variable (MAR) or an
unobserved variable (MNAR), with MCAR being benign and the latter two
biased. Football relevant examples are provided in Supplement X, and
more detailed explanations of MCAR, MAR, MNAR can be found elsewhere
[@Sainani2015; @Nakagawa2008; Sterne2009].

The three categorisations of missingness types provide a useful
framework to describe bias arising from missing data. If you strongly
suspect data is MCAR, you can be more certain your data isn't biased,
otherwise there may be bias in your sample. This does not mean data MAR
or MNAR make data invalid, instead it identifies potential biases, which
can help improve future research design. However, these categorisations
do not necessarily guide the user in their subsequent actions in
analysis, or describe specifically how values go missing. To identify
possible mechanisms for missingness, the data must be explored. It can
be challenging to identify the missingness mechanism in data, but it is
critical, since missing values could change the outcome of a study. This
could mean the work cannot be reproduced, and worse, may lead scientists
and practitioners to the wrong outcome.

It is imperative that missing data (or lack thereof --Absence of
evidence is not evidence of absence) are reported. However, outside of
sports science, researchers seldom consider missing data [@Schafer2002].
Moreover, while there has been significant interest in modelling missing
data [@refs], the exploration and reporting of missing data have received
less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
of articles that report missing data in football research; (2) highlight
several bad missing data practices; and (3) provide recommendations for
exploring and reporting missingness.

>>>>

A missing value should have been observed or measured in the data, but was not. For example, a goal scored in football takes place, but the scorekeeper wasn't paying attention and did not record an event. This "goal" is now a missing value. Compare this to a game being rained out and no game taking place. There is no missing data because there is no data to be observed. A final example - a study screens out individuals who do not meet criteria for the study. These are not missing values, because there was no intent for these to be observed or measured.

Data does not go missing the same way every time. There are three broad
categorisations describing why data can be missing: Missing Completely
at Random (MCAR), Missing at Random (MAR), and Missing Not at Random
(MNAR). These categorisations help describe whether the missingness
occurs by chance (MCAR), due to some observed variable (MAR) or an
unobserved variable (MNAR), with MCAR being the least harmful, and MAR and MNAR
biased [@Newman2014]. Football relevant examples are provided in Supplement X, and
more detailed explanations of MCAR, MAR, MNAR can be found elsewhere
[@Sainani2015; @Nakagawa2008; Sterne2009].

The three categorisations of missingness types provide a useful
framework to describe bias arising from missing data. If you strongly
suspect data is MCAR, your results are less likely to be biased. If you suspect
MAR or MNAR, there is bias in your results. This does not mean data MAR
or MNAR make results invalid, instead it identifies potential bias, which can be
used to help improve future research design. However, these categorisations
do not necessarily guide the user in their subsequent actions in
analysis, or describe specifically how values go missing. To identify
possible mechanisms for missingness, the data must be explored. It can
be challenging to identify the missingness mechanism in data, but it is
critical, since missing values could change the outcome of a study. This
could mean the work cannot be reproduced, and worse, may lead scientists
and practitioners to the wrong outcome.

It is imperative that missing data (or lack thereof --Absence of
evidence is not evidence of absence) are reported. However, outside of
sports science, researchers seldom consider missing data [@Schafer2002] (<< NOTE - I'm not sure we mean that?).
Moreover, while there has been significant interest in modelling missing
data [@refs], the exploration and reporting of missing data have received
less attention [@Schafer2002]. This article aimed to: (1) estimate the proportion
of articles that report missing data in football research; (2) highlight
several bad missing data practices; and (3) provide recommendations for
exploring and reporting missingness.
